---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r setoptions}
library(knitr)
opts_chunk$set(echo = TRUE)
options(scipen = 999)
library(ggplot2)
```

## Loading and preprocessing the data

```{r}
activity <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?

Create a histogram of the total number of steps taken each day. 

```{r, fig.height = 4, fig.width = 6}
perDay <- aggregate(steps ~ date, data = activity, FUN = "sum")
qplot(steps, data = perDay, geom = "bar", main = "Total steps taken each day", xlab = "Steps", ylab = "Number of days")
mean = mean(perDay$steps)
median = median(perDay$steps)
```

The mean number of steps per day is `r mean`, and the median is `r median`. 

## What is the average daily activity pattern?

```{r, fig.height = 4, fig.width = 6}
perInterval <- aggregate(steps ~ interval, data = activity, FUN = "mean")
qplot(interval, steps, data = perInterval, geom = "line", main = "Total steps taken each interval", xlab = "Interval", ylab = "Number of steps")
maxInterval = perInterval[perInterval$steps == max(perInterval$steps), "interval"]
```

The interval with the maximum number of steps is `r maxInterval`. 

## Imputing missing values

```{r}
numberOfNA <- sum(is.na(activity$steps))
```
There are `r numberOfNA` missing values in the data set. 

Fill in missing values in a copy of the dataset by using mean number of steps for that 5-minute interval. 
```{r}
activity2 <- activity
for (i in perInterval$interval){
    activity2[is.na(activity2$steps) & activity2$interval == i,"steps"] <- perInterval[perInterval$interval == i,"steps"] 
}
```
Create a histogram of the total number of steps taken each day (without missing values). 

```{r, fig.height = 4, fig.width = 6}
perDay2 <- aggregate(steps ~ date, data = activity2, FUN = "sum")
qplot(steps, data = perDay2, geom = "bar", main = "Total steps taken each day", xlab = "Steps", ylab = "Number of days")
mean2 = mean(perDay2$steps)
median2 = median(perDay2$steps)
```

The mean number of steps per day is `r mean2`, and the median is `r median2`. 

The mean and median are roughly the same as from the first part. The impact of inputing missing data is that days which had no data now have around the mean/median number of steps after being filled in. 

## Are there differences in activity patterns between weekdays and weekends?

```{r}
activity2$date <- as.Date(activity2$date, "%Y-%m-%d")
activity2$weekday <- "Weekday"
activity2[weekdays(activity2$date) %in% c("Saturday", "Sunday"), "weekday"] <- "Weekend"
activity2$weekday <- factor(activity2$weekday)
perInterval2 <- aggregate(steps ~ interval + weekday, data = activity2, FUN = "mean")
qplot(interval, steps, data = perInterval2, facets = weekday~., geom = "line", main = "Total steps taken each interval", xlab = "Interval", ylab = "Number of steps")
```
